---
title: "Effect of different Heat wave timing on cardiovascular mortality in France with DLNM - Sensitivity Analysis with Temperature Percentiles"
author: "Anna Alari and Noemie Letellier"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE}
library(xlsx)
library(Hmisc)
library(lubridate)
library(effects)
library(ggplot2)
library(prettyR)
#library(mgcv)
#library(jtools)
#library(broom)
library(segmented)
library(splines)
library(mgcViz)
library(ggeffects)
library(devtools)
library(zoo)
#library(grid)
library(mvmeta)
library(dlnm)
library(knitr)
library(kableExtra)
library(tidyverse)
#library(R2jags)
#library(rjags)
library("writexl")
library(metafor)
library(forestplot)
library("writexl")
library(hrbrthemes)
library(tidyverse)
library(ggridges)
library(viridis)
library(ggpubr)
library(R.utils)
library(meta)
library(jtools)
library(huxtable)
library(fmsb)

load("C:/Users/aalari/Dropbox/Noemie/FInal Analysis - SpF + CepiDC/villes_s.RData")

```


```{r}
names(villes_s)[2]<-"douai-lens"
```



## 1. FIRST STAGE ANALYSIS: Estimation of the exposure-response associations at city level for 2 periods of the summer

I decided to exclude the months of may and September, when extreme heat event never happen in France. 
Then I define two periods of the summer:  
- 1st part of the summer: 1st of june until 15th of July
- 2nd part of the summer: From 16th of July until end of August


```{r}
# Create an database for each period
summer1<-list()
summer2<-list()


for (i in 1:length(villes_s)){
  summer1[[i]]<-villes_s[[i]][which(villes_s[[i]]$day > 31 & villes_s[[i]]$day<77),]
  summer1[[i]]$time<-1:nrow(summer1[[i]])
  summer2[[i]]<-villes_s[[i]][which(villes_s[[i]]$day > 76 & villes_s[[i]]$day<124),]
  summer2[[i]]$time<-1:nrow(summer2[[i]])
  }

names(summer1)<-names(villes_s)
names(summer2)<-names(villes_s)

```


## Create a new variable for percentile of temperature


```{r}
for (i in 1: length(villes_s)){
summer1[[i]]$temp_perc <- percentile(summer1[[i]]$tempmoy)
summer2[[i]]$temp_perc <- percentile(summer2[[i]]$tempmoy)
}

lapply(summer1,function(x){summary(x$temp_perc,na.rm=TRUE)})
lapply(summer2,function(x){summary(x$temp_perc,na.rm=TRUE)})

```




#### Defintion of the cross-basis function (with its two dimensions) for the temperature of each month:

```{r}
# 1.SPECIFICATION PARAMETERS OF THE EXPOSURE-RESPONSE DIMENSION OF THE CROSS-BASIS

argvar_summer1<-list()
argvar_summer2<-list()

for (i in 1: length(villes_s)){
argvar_summer1[[i]] <- list(fun="ns", knots = quantile(summer1[[i]]$temp_perc,c(10,75,90)/100, na.rm=T),
  Bound=range(summer1[[i]]$temp_perc,na.rm=T))

argvar_summer2[[i]] <- list(fun="ns", knots = quantile(summer2[[i]]$temp_perc,c(10,75,90)/100, na.rm=T),
  Bound=range(summer2[[i]]$temp_perc,na.rm=T))
}

names(argvar_summer1)<-names(villes_s)
names(argvar_summer2)<-names(villes_s)

# 2.SPECIFICATION PARAMETERS OF THE LAG-ASSOCIATION DIMENSION OF THE CROSS-BASIS
# Definition of the maximum lag: I chose 7 days
maxlag <- 7

arglag <- list(fun="ns",knots=logknots(maxlag,nk=3))

#CREATE CROSSBASIS OBJECTS

cb_summer1<-list()
cb_summer2<-list()

for (i in 1: length(villes_s)){
cb_summer1[[i]] <- crossbasis(summer1[[i]]$temp_perc,maxlag,argvar_summer1[[i]],arglag, group=summer1[[i]]$year)  
cb_summer2[[i]] <- crossbasis(summer2[[i]]$temp_perc,maxlag,argvar_summer2[[i]],arglag, group=summer2[[i]]$year)  
}

names(cb_summer1)<-names(villes_s)
names(cb_summer2)<-names(villes_s)
```

#### Fit the models for each city and each period for CARDIOVASCULAR DEATHS
Include in the models the crossbasis term of temperature, along with the indicator for day of the week (Jours), bank holidays (Vacances) and natural cubic spline of time with 8 df per year.

```{r}
mod_summer1<-list()
mod_summer2<-list()

for (i in 1: length(villes_s)){
cb_func1<-cb_summer1[[i]]
mod_summer1[[i]] <- glm(cv_tot ~ cb_func1 + Jours + ns(day, 4) + ns(time,3), data=summer1[[i]], family=quasipoisson)
cb_func2<-cb_summer2[[i]]
mod_summer2[[i]] <- glm(cv_tot ~ cb_func2 + Jours + ns(day, 4) + ns(time,3), data=summer2[[i]], family=quasipoisson)
}

```

We get predictions for each model centering on mft value:

```{r}

mft1 <-lapply(summer1,function(x){median(x$temp_perc,na.rm=TRUE)})
mft2 <-lapply(summer2,function(x){median(x$temp_perc,na.rm=TRUE)})

pred_summer1<-list()
pred_summer2<-list()

for (i in 1: length(villes_s)){
  pred_summer1[[i]] <- crosspred(cb_func1, mod_summer1[[i]],cen=mft1[[i]], by=1)
  pred_summer2[[i]] <- crosspred(cb_func2, mod_summer2[[i]],cen=mft2[[i]], by=1)
  }
```


### PLOT - OVERALL WITH DIFFERENT FUNCTION FOR EACH PERIOD

we compare the different shapes of the cumulative exposure-response association (in terms of relative risks (RR) and centered in the mft) across the 7 days of lag, estimated for each period of the summer

```{r, echo = FALSE}

xlab <- "Percentile"

layout(matrix(c(1,2,3,4,5,6),ncol=3,nrow=2,byrow=T))
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
for (i in 1:6){
plot(pred_summer1[[i]],"overall",col="blue",axes=T,lab=c(6,5,7), xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), main=names(villes_s)[i])
lines(pred_summer2[[i]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
}

layout(matrix(c(1,2,3,4,5,6),ncol=3,nrow=2,byrow=T))
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[7]],"overall",col="blue",axes=T,lab=c(6,5,7), ylim=c(0,4), xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), main=names(villes_s)[7])
lines(pred_summer2[[7]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
for (i in 8:12){
#par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[i]],"overall",col="blue",axes=T,lab=c(6,5,7), xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), main=names(villes_s)[i])
lines(pred_summer2[[i]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
}
layout(matrix(c(1,2,3,4,5,6),ncol=3,nrow=2,byrow=T))
for (i in 13:15){
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[i]],"overall",col="blue",axes=T,lab=c(6,5,7),xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), main=names(villes_s)[i])
lines(pred_summer2[[i]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
}
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[16]],"overall",col="blue",axes=T,lab=c(6,5,7),xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), ylim=c(0, 4), main=names(villes_s)[16])
lines(pred_summer2[[16]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
for (i in 17:18){
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[i]],"overall",col="blue",axes=T,lab=c(6,5,7),xlab=xlab,
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), main=names(villes_s)[i])
lines(pred_summer2[[i]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
}

layout(matrix(c(1,2,3,4,5,6),ncol=3,nrow=2,byrow=T))
for (i in 19:21){
par(mar=c(5,4,3,1),mgp=c(3,1,0),las=1,cex.axis=0.9,cex.lab=1)
plot(pred_summer1[[i]],"overall",col="blue",axes=T,lab=c(6,5,7),xlab=xlab,
  ylab="RR", main=names(villes_s)[i], ci.arg=list(density=20,angle=-45,col=4))
lines(pred_summer2[[i]], col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period","Second Period")
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")
}
```


The set of 20 coefficients of the cross-basis in each city are then reduced. We reduce the fit of a bi-dimensional DLNM to summaries expressed by parameters of one-dimensional basis. 

```{r}
####################################################################
# BUILT OBJECTS WHERE RESULTS WILL BE STORED
#   y- IS THE MATRIX FOR THE OUTCOME PARAMETERS
#   S- IS THE LISTS OF (CO)VARIANCE MATRICES

# OVERALL CUMULATIVE SUMMARIES
yall1 <- matrix(NA,length(villes_s),4,dimnames=list(names(villes_s),paste("b",seq(4),sep="")))
yall2 <- matrix(NA,length(villes_s),4,dimnames=list(names(villes_s),paste("b",seq(4),sep="")))

# (CO)VARIANCE MATRICES
cov1 <- vector("list",length(villes_s))
cov2 <- vector("list",length(villes_s))

names(cov1)<-names(cov2) <- names(villes_s)




for (i in 1: length(villes_s)){
    cb_func1<-cb_summer1[[i]]
    red_par1 <- crossreduce(cb_func1,mod_summer1[[i]], cen=mft1[[i]])
    yall1[i,] <- coef(red_par1)  #I STORE THE RESULTS for the OVERALL CUMULATIVE SUMMARY FOR THE MAIN MODEL
    cov1[[i]] <- vcov(red_par1)  # I store the covariance matrix
    cb_func2<-cb_summer2[[i]]
    red_par2 <- crossreduce(cb_func2,mod_summer2[[i]], cen=mft2[[i]])
    yall2[i,] <- coef(red_par2)
    cov2[[i]] <- vcov(red_par2)
    }




```

For each city we derive the vector with 4 reduced parameters of the natural spline of temperature for the overall cumulative summary association. 


## 2. Second stage analysis to obtain pooled estimates

In the second stage analysis, we: 
- RUN THE MULTIVARIATE META-ANALYTICAL MODELS WITH mvmeta
- CREATE BASIS VARIABLES USING onebasis, TO BE USED FOR PREDICTION
- OBTAIN PREDICTIONS THROUGH crosspred (dlnm)

```{r}
library(mvmeta)

# OVERALL CUMULATIVE SUMMARY FOR THE MAIN MODEL
mvall1 <- mvmeta(yall1,cov1, method="ml")
mvall2 <- mvmeta(yall2,cov2,method="ml")

summary(mvall1)
summary(mvall2)
```


```{r}
# CREATE BASES FOR PREDICTION

# BASES OF TEMPERATURE AND LAG USED TO PREDICT, EQUAL TO THAT USED FOR ESTIMATION
# COMPUTED USING THE ATTRIBUTES OF THE CROSS-BASIS USED IN ESTIMATION

x1<-lapply(summer1,function(x){x[,"temp_perc"]})
tempall1<-rbind(x1[["avignon"]],x1[["douai-lens"]],x1[["bordeaux"]],x1[["clermont"]],x1[["dijon"]],x1[["grenoble"]], x1[["lehavre"]], x1[["lille"]],x1[["lyon"]],x1[["marseille"]], x1[["montpellier"]],x1[["nancy"]],x1[["nantes"]],x1[["nice"]],x1[["paris"]],x1[["rennes"]],x1[["rouen"]],x1[["saint_etienne"]],x1[["strasbourg"]],x1[["toulon"]],x1[["toulouse"]])

x2<-lapply(summer2,function(x){x[,"temp_perc"]})
tempall2<-rbind(x2[["avignon"]],x2[["douai-lens"]],x2[["bordeaux"]],x2[["clermont"]],x2[["dijon"]],x2[["grenoble"]], x2[["lehavre"]], x2[["lille"]],x2[["lyon"]],x2[["marseille"]], x2[["montpellier"]],x2[["nancy"]],x2[["nantes"]],x2[["nice"]],x2[["paris"]],x2[["rennes"]],x2[["rouen"]],x2[["saint_etienne"]],x2[["strasbourg"]],x2[["toulon"]],x2[["toulouse"]])

# minimum<-min(data.frame(lapply(villes_s,function(x){min(x$tempmoy, na.rm=TRUE)})))
# maximum<-max(data.frame(lapply(villes_s,function(x){max(x$tempmoy, na.rm=TRUE)})))

xvar1 <- seq(min(tempall1, na.rm=TRUE),max(tempall1, na.rm=TRUE),by=0.1)
xvar2 <- seq(min(tempall2, na.rm=TRUE),max(tempall2, na.rm=TRUE),by=0.1)

bvar1 <- onebasis(xvar1, "ns", knots = quantile(tempall1,c(10,75,90)/100, na.rm=T),
  Bound=range(tempall1,na.rm=T))
bvar2 <- onebasis(xvar2, "ns", knots = quantile(tempall2,c(10,75,90)/100, na.rm=T),
  Bound=range(tempall2,na.rm=T))

#bvar <- do.call("onebasis",c(list(x=xvar),attr(cb,"argvar")))

# xlag <- 1:7
# blag <- onebasis(xlag, fun="ns",knots=logknots(maxlag,nk=3))


####################################################################
# City-SPECIFIC FIRST-STAGE SUMMARIES

cityall1 <- apply(yall1,1,function(x) exp(bvar1%*%x))
cityall2 <- apply(yall2,1,function(x) exp(bvar2%*%x))

####################################################################
```


```{r}
# PREDICTION FOR A GRID OF TEMPERATURE AND LAG VALUES

# OVERALL CUMULATIVE SUMMARY ASSOCIATION FOR MAIN MODEL

model.class <- class(mvall1)
# coeff <- dlnm:::getcoef(mvall2,model.class)
# vcova <- dlnm:::getvcov(mvall2,model.class)
#model.link <- dlnm:::getlink(mvall1,model.class)

# test<-c(coef(mvall1)[[1]], coef(mvall1)[[2]], coef(mvall1)[[3]], coef(mvall1)[[4]])
# vcova<-matrix(ncol=4, nrow=4, dlnm:::getvcov(mvall1,model.class))

cpall1 <- crosspred(bvar1, coef=dlnm:::getcoef(mvall1,model.class), vcov=dlnm:::getvcov(mvall1,model.class), model.link="log",by=0.1,from=min(tempall1, na.rm=TRUE),to=max(tempall1, na.rm=TRUE), cen=median(tempall1, na.rm=TRUE))

cpall2 <- crosspred(bvar2, coef=dlnm:::getcoef(mvall2,model.class), vcov=dlnm:::getvcov(mvall2,model.class), model.link="log",by=0.1,from=min(tempall2, na.rm=TRUE),to=max(tempall2, na.rm=TRUE), cen=median(tempall2, na.rm=TRUE))

```

```{r}
plot(cpall1,col="blue",axes=T,lab=c(6,5,7),xlab="Percentile of Temperature",
  ylab="RR", ci.arg=list(density=20,angle=-45,col=4), ylim=c(0.80,1.4))
lines(cpall2, col="red", ci="area", ci.arg=list(density=20,col=2))
legend("topleft",c("First Period \n (I²=1.0%, X² test p-val = 0.4943)","Second Period \n (I²=76.5%, X² test p-val = 0.0000)"), horiz = TRUE
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=0.7,ncol=1,inset=0.01,bty="n")

```


```{r, echo=FALSE}
x1<-lapply(summer1,function(x){x[,c("temp_perc")]})
x2<-lapply(summer2,function(x){x[,c("temp_perc")]})

# x1<-list()
# x2<-list()

# for (i in 1:length(villes_s)){
#   x1[[i]]<-x_1[[i]][which(x_1[[i]]$day > 31 & x_1[[i]]$day<77),]
#   x2[[i]]<-x_2[[i]][which(x_2[[i]]$day > 76 & x_2[[i]]$day<124),]
#   }

names(x1)<-names(villes_s)
names(x2)<-names(villes_s)

tempall1<-rbind(x1[["avignon"]],x1[["douai-lens"]],x1[["bordeaux"]],x1[["clermont"]],x1[["dijon"]],x1[["grenoble"]], x1[["lehavre"]], x1[["lille"]],x1[["lyon"]],x1[["marseille"]], x1[["montpellier"]],x1[["nancy"]],x1[["nantes"]],x1[["nice"]],x1[["paris"]],x1[["rennes"]],x1[["rouen"]],x1[["saint_etienne"]],x1[["strasbourg"]],x1[["toulon"]],x1[["toulouse"]])

tempall2<-rbind(x2[["avignon"]],x2[["douai-lens"]],x2[["bordeaux"]],x2[["clermont"]],x2[["dijon"]],x2[["grenoble"]], x2[["lehavre"]], x2[["lille"]],x2[["lyon"]],x2[["marseille"]], x2[["montpellier"]],x2[["nancy"]],x2[["nantes"]],x2[["nice"]],x2[["paris"]],x2[["rennes"]],x2[["rouen"]],x2[["saint_etienne"]],x2[["strasbourg"]],x2[["toulon"]],x2[["toulouse"]])
```


```{r, echo=FALSE, fig.dim=c(8,8)}
layout(matrix(1:4,ncol=1),heights=c(0.8,rep(0.20,2),0.20))
par(mar=c(0.2,5,2,1),las=1,cex.axis=0.8,cex.lab=1)
plot(cpall1,col="blue",lab=c(6,5,7),xlab="Percentile of Temperature",
  ylab="RR", axes=F, ci.arg=list(density=20,angle=-45,col=4), xlim=c(min(xvar1),max(xvar1)), ylim=c(0.80,1.4))
lines(cpall2, col="red", ci="area", ci.arg=list(density=20,col=2), xlim=c(min(xvar2),max(xvar2)))
legend("topleft",c("First Period \n (I²=1.0%, X² test p-val = 0.4943)","Second Period \n (I²=76.5%, X² test p-val = 0.0000)"), horiz = TRUE
  ,xpd = TRUE,col=c("blue","red"),lwd=2,bg="white"
  ,cex=1,ncol=1,inset=0.01,bty="n")
# abline(v=mft[i],lty=1, col="blue")
# abline(v=quantile(villes_s[[i]]$tempmoy,c(0.1,0.75,0.90)),lty=2)
axis(2,at=c(0,round(max(cpall1$allRRfit)*(1/8),digits =2),round(max(cpall1$allRRfit)*(2/8),digits=2),round(max(cpall1$allRRfit)*(3/8),digits = 2),round(max(cpall1$allRRfit)*(4/8),digits=2),round(max(cpall1$allRRfit)*(5/8),digits=2), round(max(cpall1$allRRfit)*(6/8),digits=2), round(max(cpall1$allRRfit)*(7/8),digits=2),round(max(cpall1$allRRfit),digits=2),2))
par(mar=c(0.1,5,0.3,1))
hist(tempall1,axes=F,col=grey(0.95),ylab="Temp Freq\nFirst Period",xlim=c(min(xvar1),max(xvar1)),breaks=20, main="")
# abline(v=mft[i],lty=1, col="blue")
# abline(v=quantile(villes_s[[i]]$tempmoy,c(0.1,0.75,0.90)),lty=2)
axis(2,at=0:50*100)
par(mar=c(0.1,5,0.2,1))
hist(tempall2,axes=F,col=grey(0.95),ylab="Temp Freq\nSecond Period",xlim=c(min(xvar2),max(xvar2)), breaks=20, main="")
axis(2,at=0:50*100)
par(mar=c(4,1,1,1))
plot(cpall2$predvar,rep(0,length(cpall2$predvar)),type="n",yaxt="n",ylab="",xlim=c(min(xvar1),max(xvar1)),
     xlab="Percentile of Temperature",frame.plot=F)



```


